Learning a functional distance between organizations with unsupervised embeddings of scientific mobility

### 2020-07-16 paper/current/main.tex
typo

Human mobility and migration drives major social phenomenon such as the growth and evolution of cities~\autocite{wef2017migration, curiel2018citygravity},  epidemics~\autocite{kraemer2020covid, truscott2012epidemicgravity, xia2005measlesgravity}, economies~\autocite{kaluza2010cargo, kerr2011immigration}, innovation~\autocite{kaiser2018innovation, armano2017innovation},  and scientific research~\autocite{sugimoto2017mostimpact, petersen2018multiscale, franzoni2014advantage, morgan2018prestige, rodrigues2016mobility}. 
Common models of mobility, such as the Gravity Model~\autocite{zipf1946gravity, xia2005measlesgravity, jung2008highwaygravity, barbosa2018human, curiel2018citygravity, lewer2008immigrationgravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity}, often successfully capture mobility between places using their population and the \textit{distance} between them.
However, geographic distance is less relevant in the increasingly-globalized world~\autocite{czaika2018migration, czaika2018globalisation, buchan2009globalization} where the physical distance barrier shrinks while the impact of linguistic, economic, cultural, and historical similarity increases.
Alternative notions of functional distance~\autocite{brown1970migration, brown1970functional, kim2018functional} attempt to re-define distance by using some of the aforementioned factor~\autocite{boschma2005proximity,  brown1970functional, brown1970migration, kim2018functional} but suffer from practical and theoretical challenges. 
Here, using a rich dataset of scientific mobility, we demonstrate that a data-driven neural embedding approach~\autocite{mikolov2013word2vec} can successfully encodes a functional distance between organizations and captures the latent and multi-faceted structure of scientist's mobility.
We demonstrate that the cosine distance between organization embedding vectors, constructed using the affiliation trajectories of scientists in the Web of Science database, better explain and predict real mobility than geographic distance. 
Exploiting their unique properties, we interpret the embedding space and observe how it captures latent factors related to culture, language, and prestige of organizations at multiple levels of granularity.
The ability yo learn scalable, dense, and meaningful vector representations of mobility directly from the data can open up a new avenue of studying human mobility.

>>>>

Human mobility and migration drives major social phenomenon such as the growth and evolution of cities~\autocite{wef2017migration, curiel2018citygravity},  epidemics~\autocite{kraemer2020covid, truscott2012epidemicgravity, xia2005measlesgravity}, economies~\autocite{kaluza2010cargo, kerr2011immigration}, innovation~\autocite{kaiser2018innovation, armano2017innovation},  and scientific research~\autocite{sugimoto2017mostimpact, petersen2018multiscale, franzoni2014advantage, morgan2018prestige, rodrigues2016mobility}. 
Common models of mobility, such as the Gravity Model~\autocite{zipf1946gravity, xia2005measlesgravity, jung2008highwaygravity, barbosa2018human, curiel2018citygravity, lewer2008immigrationgravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity}, often successfully capture mobility between places using their population and the \textit{distance} between them.
However, geographic distance is less relevant in the increasingly-globalized world~\autocite{czaika2018migration, czaika2018globalisation, buchan2009globalization} where the physical distance barrier shrinks while the impact of linguistic, economic, cultural, and historical similarity increases.
Alternative notions of functional distance~\autocite{brown1970migration, brown1970functional, kim2018functional} attempt to re-define distance by using some of the aforementioned factor~\autocite{boschma2005proximity,  brown1970functional, brown1970migration, kim2018functional} but suffer from practical and theoretical challenges. 
Here, using a rich dataset of scientific mobility, we demonstrate that a data-driven neural embedding approach~\autocite{mikolov2013word2vec} can successfully encodes a functional distance between organizations and captures the latent and multi-faceted structure of scientist's mobility.
We demonstrate that the cosine distance between organization embedding vectors, constructed using the affiliation trajectories of scientists in the Web of Science database, better explain and predict real mobility than geographic distance. 
Exploiting their unique properties, we interpret the embedding space and observe how it captures latent factors related to culture, language, and prestige of organizations at multiple levels of granularity.
The ability to learn scalable, dense, and meaningful vector representations of mobility directly from the data can open up a new avenue of studying human mobility.
